---
title: "A1_gapminder"
author: "Laura Conover"
output: html_document
---
```{r echo=FALSE}
# install.packages('gapminder')
# install.packages('DT')
```

```{r echo=FALSE}
library(gapminder)
library(DT)
```

  
The "gapminder" data set includes data for `r nrow(gapminder)` countries on variables such as life expectancy (average = `r round(mean(gapminder$lifeExp),digits=2)` years), population size (average = `r round(mean(gapminder$pop),digits=0)` people), and per-capita GDP (average = \$`r round(mean(gapminder$gdpPercap),digits=2)`).  Yes, assumptions were made about units.

The plots below display some interesting relationships between these given variables.  
  

```{r echo=FALSE}
plot(log(gapminder$pop),gapminder$lifeExp, type = "p", main = "Relationship between total population and life expectancy", xlab = "Log(Population)", ylab = "Life Expectancy")
plot(log(gapminder$gdpPercap),gapminder$lifeExp, type = "p", main = "Relationship between per-capita GDP and life expectancy", xlab = "Log(GDP)", ylab = "Life Expectancy")
```

Presented below is a searchable form of the dataset used.  

```{r echo=FALSE}
DT::datatable(gapminder)
```