---
title: "Final project"
author: "Laura Conover"
output_format: html_document
output:
  html_document:
    code_folding: hide
    keep_md: true
  word_document: default
knit: (function(inputFile, encoding) { 
  rmarkdown::render(inputFile, 
                    encoding = encoding, 
                    output_format = rmarkdown::yaml_front_matter(inputFile)$output_format,
                    output_dir = here::here("output",
                      rmarkdown::yaml_front_matter(inputFile)$output_format
                      )
                    )
  })
---

<!-- See above with the 'output_format' and 'knit' arguments for how to output
     your knitted document to a subfolder in `output` instead of in `markdown` -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### I'm not entirely sure, but I feel like this document is now mostly unnecessary, as the vast majority of the needed information is now in the "writeup" doc.  But I also do not want to delete it just in case.

Required packages:
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(svglite))
```

Folder set-up if required:
```{r}
#source(here::here("scripts","folder_setup.R"))
```

Download data from the Harvard dataverse and load into R:
```{r}
source(here::here("scripts", "load_data.R"))
```

_A description of the data can be found in this project's README file._
_These datasets are somewhat messy.  Participant ids were reused at different sites, and stimulus files were renamed between analyses.  Additionally, there are a number of variables that are not relevant to the current project._

Create new dataset with only the relevant variables and save it as a file:
```{r}
source(here::here("scripts", "clean_data.R"))
```

Display data.
```{r}
DT::datatable(by_subject)
DT::datatable(by_laugh)
```

Create and save plots.

```{r}
source(here::here("scripts", "plot_data.R"))

laugh_plot
subject_plot
```


```{r, echo=TRUE}
devtools::session_info()
```

