---
title: 'Data Science and Programming Assignment 04: Tidy data and joins'
author: "Laura Conover"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```


## Objective

The goal of this homework is to solidify your data wrangling skills by working 
some realistic problems in the grey area between data aggregation and data 
reshaping.

Load required packages:
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```

## Exercise 1: Univariate Data Reshaping (30%)

### Univariate Option 1

_You are expected to use `pivot_wider()` and `pivot_longer()` for reshaping, and `ggplot2` for the plot_
  
_1. Make a tibble with one row per year, and columns for life expectancy for two or more countries_

Life Expectancy for Ireland and the UK from 1952 to 2007
``` {r}
ex1 <- gapminder %>% 
   filter(country %in% c('Ireland', 'United Kingdom')) %>% 
   pivot_wider(id_cols = year, names_from = country, values_from = lifeExp) %>%
   print
```
_2. Take advantage of this new data shape to scatterplot life expectancy for one country against that of another._

```{r}
ex1 %>% 
   ggplot(aes(Ireland,`United Kingdom`)) +
   geom_point() +
   geom_smooth(method = 'lm') +
   ggtitle("Correlated Life Expectancy in Ireland and the UK from 1952 to 2007")
```

_3. Re-lengthen the data._
```{r}
ex1 %>% 
   pivot_longer(cols = c('Ireland', 'United Kingdom'), names_to = 'country', values_to = 'life expectancy') %>% 
   DT::datatable()
```

## Exercise 2: Multivariate Data Reshaping (30%)

_You are expected to use `pivot_wider()` and `pivot_longer()` for reshaping._

### Multivariate Option 1

*1. Make a tibble with _one row per year_, and _columns_ for life expectancy _and_ GDP per capita for two or more countries.*

Life Expectancy and GDP for Ireland and the UK from 1952 to 2007
```{r}
ex2 <- gapminder %>% 
   filter(country %in% c('Ireland', 'United Kingdom')) %>% 
   pivot_wider(id_cols = year, names_from = country, values_from = c(lifeExp,gdpPercap)) %>%
   print
```

_2. Print this table in a pretty format, using either `knitr::kable()` or `DT::datatable()` depending on the number of rows of data._

```{r}
DT::datatable(ex2)
```

_3. Re-lengthen the data._
```{r}
ex2 %>% 
   pivot_longer(cols = -year, names_to = c(".value","country"), names_sep = '_') %>% 
   DT::datatable()
```


## Exercise 3: Table Joins (30%)

_Read in the made-up wedding guestlist and email addresses using the following lines.  Then, complete the following tasks using the tidyverse (`tidyr`, `dplyr`, ...)._

``` {r}
guest <- read_csv("https://raw.githubusercontent.com/USF-Psych-DataSci/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/USF-Psych-DataSci/Classroom/master/data/wedding/emails.csv")
```


_1.For each guest in the guestlist (`guest` tibble), add a column for email address, which can be found in the `email` tibble._

List each person individually by name in the 'email' tibble:
```{r}
email_sep <- email %>% 
   separate_rows(guest, sep = ", ") %>% 
   rename(name = guest)
```

Combine the two tibbles:
```{r}
guest %>% 
   left_join(email_sep, by = 'name') %>% 
   DT::datatable()
```

_2.Who do we have emails for, yet are not on the guestlist?_
``` {r}
email_sep %>% 
   anti_join(guest, by = 'name') %>% 
   knitr::kable()
```


_3.Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist)._
```{r}
guest %>% 
   full_join(email_sep, by = 'name') %>% 
   DT::datatable()
#note that this list is not saved (since this is the last step in the assignment).  In a practical situation, it would likely be assigned to a new variable
```