---
title: "Data Science and Programming Assignment 03: Write a Report on Exploring a Dataset"
author: Laura Conover
output: 
  html_document:
    keep_md: true
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```

__Due February 17, 2020 at 23:59__


## Overview

The goal is to manipulate and explore a dataset with the `dplyr` package, 
complemented by visualizations made with `ggplot2`. You can think of this as a 
deeper version of Homework 02.

## Assignment

Pick _*three*_ of the tasks below, and, for each selected task, produce: 

- a tibble, using `dplyr` as your data manipulation tool;
- an accompanying plot of data from the tibble, using `ggplot2` as your visualization tool; and
- some dialogue about what your tables/figures show (doesn't have to be much).
  - The idea is to structure your assignment like it's an actual report someone
    might read.

---

### Task Option 3

Work with `gapminder`. Look at the spread/variability of GDP per capita within 
the continents.
  
_Tibble:_
  
```{r echo=FALSE}
gapminder %>% 
  group_by(continent) %>% 
  summarise(Minimum = min(round(gdpPercap, digits =2)),
            Maximum = max(round(gdpPercap, digits =2)),
            Range = Maximum-Minimum) %>%
  knitr::kable()
```

_Plot:_
  
```{r echo=FALSE}
gapminder %>% 
  ggplot(aes(gdpPercap)) +
  geom_density() +
  facet_wrap(~ continent, nrow = 2, scales = "fixed") 
```

<!-- Plotting GDP on a log scale might be better -->

_Dialogue:_
  
As you can see in the graphs above, every continent has a right-tailed distribution of GDP per capita.  Although most continents have a range across the years between 20,000 and 50,000, Asia has the greatest spread with a range over 113,000.

  
### Task Option 7

Work with `psych::bfi`. Compute composite scores for the five Big Five scores, 
including both person-mean scores and person-sd scores. Describe the 
distributions of the different personality variables.
 
_Tibble:_
```{r echo=TRUE}
composite_bfi<-psych::bfi %>% 
  rownames_to_column(var = "id") %>% 
  as_tibble() %>%
  group_by(id) %>% # rowwise() will also work
  mutate(A_mean = mean(c(A1, A2, A3, A4, A5),na.rm = TRUE), 
         A_sd = sd(c(A1, A2, A3, A4, A5),na.rm = TRUE), 
         C_mean = mean(c(C1, C2, C3, C4, C5),na.rm = TRUE), 
         C_sd = sd(c(C1, C2, C3, C4, C5),na.rm = TRUE), 
         E_mean = mean(c(E1, E2, E3, E4, E5),na.rm = TRUE), 
         E_sd = sd(c(E1, E2, E3, E4, E5),na.rm = TRUE), 
         N_mean = mean(c(N1, N2, N3, N4, N5),na.rm = TRUE), 
         N_sd = sd(c(N1, N2, N3, N4, N5),na.rm = TRUE), 
         O_mean = mean(c(O1, O2, O3, O4, O5),na.rm = TRUE), 
         O_sd = sd(c(O1, O2, O3, O4, O5),na.rm = TRUE)) %>% # Best to break this up into multiple lines like I did here.
  ungroup() %>% 
  select(id,gender,education,age,A_mean,A_sd,C_mean,C_sd,E_mean,E_sd,N_mean,N_sd,O_mean,O_sd) %>% 
  print()
```

_Plot:_
```{r echo=FALSE}
composite_bfi %>% 
  ggplot() +
  geom_density(aes(A_mean), color = "red") +
  geom_density(aes(C_mean), color = "darkblue") +
  geom_density(aes(E_mean), color = "darkgreen") +
  geom_density(aes(N_mean), color = "purple") +
  geom_density(aes(O_mean), color= "goldenrod") +
  xlab("Composite Score") +
  ylab("Response Density") +
  ggtitle("Average Scores on Big 5 Traits") +
  theme_bw()
#if you put the name of the color you want in quotes, it won't make a legend.
# No, that's not the issue here. The problem is that you have color inside aes().
# Only aesthetics that reference a variable in data should be inside aes().
# Other aesthetics that are just specified as constants, like here, should be outside
# aes().
```

_Dialogue:_
  
Although individuals' average scores of three of the Big 5 traits (Openness, Extraversion, and Conscientiousness) have similar peaks and ranges, the average scores for Neuroticism form a much flatter distribution.
  
### Task Option 9

Work with `psych::bfi`. Using the composite scores from Task 7, explore how age
is related to the different personality variables. Does personality appear to 
change for different ages?
  
_Tibble:_
  
```{r echo=FALSE}
composite_bfi %>% 
  select(id,age,A_mean,A_sd,C_mean,C_sd,E_mean,E_sd,N_mean,N_sd,O_mean,O_sd) %>% 
  print()
```
 
_Plot:_
 
```{r echo=FALSE}
composite_bfi %>% 
  ggplot() +
  geom_point(aes(age, A_mean), color = "red") +
  ylab("Agreeableness Average Score") +
  xlab("Age")
composite_bfi %>% 
  ggplot() +
  geom_point(aes(age, C_mean), color = "yellow") +
  ylab("Conscientiousness Average Score") +
  xlab("Age")
composite_bfi %>% 
  ggplot() +
  geom_point(aes(age, E_mean), color = "green") +
  ylab("Extraversion Average Score") +
  xlab("Age")
composite_bfi %>% 
  ggplot() +
  geom_point(aes(age, N_mean), color = "blue") +
  ylab("Neuroticism Average Score") +
  xlab("Age")
composite_bfi %>% 
  ggplot() +
  geom_point(aes(age, O_mean), color = "purple") +
  ylab("Openness Average Score") +
  xlab("Age")
```

_Dialogue:_
  
As can be seen in the plots above, age is uncorrelated with average score on any of the Big 5 personality traits.
  
## Optional Exercise (2%)

Get a table and figure to show side-by-side. 
[This gist](https://gist.github.com/jennybc/e9e9aba6ba18c72cec26) might get you 
started.
